---
title: "Covid Figures and Maps"
author: "Steven Rashin"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F,cache = F, message = F)
library(usmap)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
library(tidyr)
library(lubridate)
library(tigris)
library(dslabs)
library(grid)
library(gridExtra)
library(cowplot)
library(plotly)
library(kableExtra)
library(zoo)

setwd("/Users/stevenrashin/Dropbox/covid data/")

### CDC linked data ###
CDC_cases <- fread('https://usafactsstatic.blob.core.windows.net/public/data/covid-19/covid_confirmed_usafacts.csv')
CDC_deaths <- fread('https://usafactsstatic.blob.core.windows.net/public/data/covid-19/covid_deaths_usafacts.csv')

### Population Data ###
population_data <- fread("https://usafactsstatic.blob.core.windows.net/public/data/covid-19/covid_county_population_usafacts.csv")

population_data_by_state <- population_data %>%
  group_by(State) %>%
  summarize(population = sum(population,na.rm = T))

#CDC_deaths_by_race_and_state <- fread('https://data.cdc.gov/api/views/ks3g-spdg/rows.csv?accessType=DOWNLOAD')

### Reshape data to long, tidyreshape is easy so use that! ###
long_cases <- pivot_longer(CDC_cases, cols=5:length(CDC_cases), names_to = "Date", values_to = "Cases")
long_cases$Date <- lubridate::mdy(long_cases$Date)

long_deaths <- pivot_longer(CDC_deaths, cols=5:length(CDC_deaths), names_to = "Date", values_to = "Deaths")
long_deaths$Date <- lubridate::mdy(long_deaths$Date)
long_deaths$`County Name` <- NULL

#check for duplicates, none found but worth checking
# any_dupes_CDC_cases <- c(paste0(long_cases$countyFIPS, long_cases$State, long_cases$stateFIPS, long_cases$Date))
# which(duplicated(any_dupes_CDC_cases))
# 
# any_dupes_CDC_deaths <- c(paste0(long_deaths$countyFIPS, long_deaths$State, long_deaths$stateFIPS, long_deaths$Date))
# which(duplicated(any_dupes_CDC_deaths))

CDC_total <- merge(long_cases, long_deaths, by = c("countyFIPS" ,"State","stateFIPS","Date"))

CDC_total <- CDC_total %>% 
  mutate(Week = lubridate::week(Date),
         Day = lubridate::yday(Date),
         Month = lubridate::month(Date))

rm(CDC_cases, CDC_deaths, long_cases, long_deaths)


### COVID tracking data from https://covidtracking.com/data/download ###
#confirmed_cases_covid_tracker <- fread("https://covidtracking.com/api/v1/us/daily.csv")
covid_tracking_data <- fread("https://covidtracking.com/api/v1/states/daily.csv")

#Rename variables as necessary to make it easy to compare between data sets
covid_tracking_data$Date <- lubridate::ymd(covid_tracking_data$date)
covid_tracking_data$State <- covid_tracking_data$state

#### Plot New Cases BY STATE with week long kernal smoother ####
# trend line modified from https://rafalab.github.io/dsbook/smoothing.html
# note that kernl = normal prodices better plots than kernel = box as some states (e.g. MS) don't produce data
# for days at a time and normal smooths this out more than box
# also, if you want to do loess, use the following code loess(new_cases ~ Day, degree=1, span = 14)$y
# for a one week average, the normal kernel does better than a loess

CDC_w_trendline <- CDC_total %>% 
  #filter(State == "MA") %>%
  group_by(State, stateFIPS, Date, Day, Week) %>%
  arrange(State, Date, countyFIPS) %>%
  summarise(total_cases_by_day = sum(Cases),
            total_deaths_by_day = sum(Deaths)) %>%
  ungroup() %>% 
  group_by(State) %>%
  mutate(new_cases = total_cases_by_day - dplyr::lag(total_cases_by_day, default = 0),
         new_deaths = total_deaths_by_day - dplyr::lag(total_deaths_by_day, default = 0)) %>%
  mutate( smooth_new_cases = ksmooth(Day, new_cases, kernel = "normal", bandwidth = 7)$y,
          smooth_total_cases = ksmooth(Day, total_cases_by_day, kernel = "normal", bandwidth = 7)$y,
          smooth_deaths = ksmooth(Day, new_deaths, kernel = "normal", bandwidth = 7)$y,
          smooth_total_deaths = ksmooth(Day, total_deaths_by_day, kernel = "normal", bandwidth = 7)$y)  %>%
  ungroup() 

### County data for headline numbers below: 
county_with_most <- CDC_total %>% 
  arrange(State, Date, countyFIPS) %>%
  group_by(countyFIPS) %>%
  mutate(new_cases = Cases - dplyr::lag(Cases, default = 0),
         new_deaths = Deaths - dplyr::lag(Deaths, default = 0)) %>%
  ungroup() %>%
  filter(Date == max(Date, na.rm = T)) %>% 
  filter(new_cases == max(new_cases))

state_with_most <- CDC_w_trendline %>%   
  filter(Date == max(Date, na.rm = T)) %>% 
  filter(new_cases == max(new_cases)) %>%
  select(State, new_cases,Date)


covid_tracking_data_w_trendline <- covid_tracking_data %>% filter(State %in% state.abb) %>%
  mutate(Week = lubridate::week(Date),
         Day = lubridate::yday(Date)) %>%
  arrange(State, Date)  %>% 
  group_by(State) %>%
  mutate( smooth_new_cases = ksmooth(Day, positiveIncrease, kernel = "normal", bandwidth = 7)$y,
          smooth_total_cases = ksmooth(Day, positive, kernel = "normal", bandwidth = 7)$y,
          smooth_deaths = ksmooth(Day, deathIncrease, kernel = "normal", bandwidth = 7)$y,
          smooth_total_deaths = ksmooth(Day, death, kernel = "normal", bandwidth = 7)$y) %>%
  ungroup() 


NE.abrv <- c("CT","ME","MA","NH","RI","VT","NJ","NY","PA")
MW.abrv <- c("IN","IL","MI","OH","WI","IA","KS","MN","MO","NE",
             "ND","SD")
S.abrv <- c("DE","DC","FL","GA","MD","NC","SC","VA","WV","AL",
            "KY","MS","TN","AR","LA","OK","TX")
W.abrv <- c("AZ","CO","ID","NM","MT","UT","NV","WY","AK","CA",
            "HI","OR","WA")
region.list <- list(
  Northeast=NE.abrv,
  Midwest=MW.abrv,
  South=S.abrv,
  West=W.abrv)

CDC_w_trendline$Region <- sapply(CDC_w_trendline$State, function(x) names(region.list)[grep(x,region.list)])

### Get global cases for headlines 

try_url <- function(url){
      tryCatch(
        expr = {
          local <- fread(url)
          return(local)
        },
        error = function(e){
            message('"ERROR Out of Date Data / Using Last Saved DATA"')
            global <- read.csv("/Users/stevenrashin/Documents/GitHub/Covid-Tracker/WHO-COVID-19-global-data.csv",stringsAsFactors = F)
            return(local)
        }
    )
}

global <- try_url("https://covid19.who.int/WHO-COVID-19-global-data.csv")
global$Date <- lubridate::ymd(gsub(pattern = "T.{1,}$", replacement = "", x = global$Date_reported))


try_url <- function(url){
      tryCatch(
        expr = {
          local <- fread(url)
          return(local)
        },
        error = function(e){
            message('"ERROR Out of Date Data / Using Last Saved DATA"')
            local <- read.csv("/Users/stevenrashin/Documents/GitHub/Covid-Tracker/daily-c19-case-counts.csv",stringsAsFactors = F)
            return(local)
        }
    )
}

local <- try_url("https://somerville-dashboardcovid.trial.opendatasoft.com/explore/dataset/daily-c19-case-counts/download/?format=csv&timezone=America/New_York&lang=en&use_labels_for_header=true&csv_separator=%2C")

local$Date <- lubridate::ymd(gsub(pattern = "T.{1,}$", replacement = "", x = local$Date))

Somerville_MA <- local %>% 
    mutate(new_cases = Positive - dplyr::lag(Positive, default = 0),
         new_deaths = Fatalities - dplyr::lag(Fatalities, default = 0),
         Day = lubridate::yday(Date),
         Week = lubridate::week(Date)) %>%
    mutate( smooth_new_cases = ksmooth(Day, new_cases, kernel = "normal", bandwidth = 7)$y) %>%
  filter(Day %in% (max(Day)-14):max(Day))

# in `r county_with_most %>% select(State) %>% pull()` with `r formatC(county_with_most %>% select(new_cases) %>% pull(), format="f", big.mark=",", digits=0)` new cases
```

## Headline Numbers 
As of the latest CDC data release on `r format(max(CDC_w_trendline$Date), '%B %d, %Y')` there are:  

* `r formatC(CDC_w_trendline %>% filter(Date == max(Date)) %>% ungroup() %>% summarize(sum(new_cases)) %>% pull(), format="f", big.mark=",", digits=0)` new cases
  + The state with the most cases is `r state_with_most %>% select(State) %>% pull()` with `r formatC(state_with_most %>% select(new_cases) %>% pull(), format="f", big.mark=",", digits=0)`
    - The county with the most cases is `r county_with_most[,5]` in `r county_with_most[,2]` with `r formatC(county_with_most %>% select(new_cases) %>% pull(), format="f", big.mark=",", digits=0)` new cases 
  + The record number of national cases per day is `r formatC(CDC_w_trendline %>% group_by(Date) %>% summarize(new_cases = sum(new_cases)) %>% filter(new_cases == max(new_cases)) %>% select(new_cases) %>% pull(), format="f", big.mark=",", digits=0)` on `r format(CDC_w_trendline %>% group_by(Date) %>% summarize(new_cases = sum(new_cases)) %>% filter(new_cases == max(new_cases)) %>% select(Date) %>% pull(), '%B %d, %Y')`
* `r formatC(CDC_w_trendline %>% filter(Date == max(Date)) %>% ungroup() %>% summarize(sum(new_deaths)) %>% pull(), format="f", big.mark=",", digits=0)` new deaths
  +  The record number of deaths per day is `r formatC(CDC_w_trendline %>% group_by(Date) %>% summarize(new_deaths = sum(new_deaths)) %>% filter(new_deaths == max(new_deaths)) %>% select(new_deaths) %>% pull(), format="f", big.mark=",", digits=0)` on `r format(CDC_w_trendline %>% group_by(Date) %>% summarize(new_deaths = sum(new_deaths)) %>% filter(new_deaths == max(new_deaths)) %>% select(Date) %>% pull(), '%B %d, %Y')`
  
Where I live, in Somerville, MA, as of `r format(max(Somerville_MA$Date), '%B %d, %Y')` there are:  

* `r Somerville_MA %>% filter(Date == max(Date)) %>% select(new_cases) %>% pull()` new cases  
* `r Somerville_MA %>% filter(Date %in% (max(Date)-14):max(Date)) %>% summarize(new_cases = sum(new_cases, na.rm = T)) %>% pull()` cases in the last two weeks  
  + Out of `r Somerville_MA %>% filter(Date == max(Date)) %>% select(Positive) %>% pull()` cases overall   
    - `r Somerville_MA %>% filter(Date == max(Date)) %>% select(Fatalities) %>% pull()` Fatalities and
    - `r Somerville_MA %>% filter(Date == max(Date)) %>% select(Recovered) %>% pull()` Recoveries

Consequently, in the United States, there are now:   

* `r formatC(CDC_w_trendline %>% ungroup() %>% summarize(sum(new_cases)) %>% pull(), format="f", big.mark=",", digits=0)` total cases  
* `r formatC(CDC_w_trendline %>% ungroup() %>% summarize(sum(new_deaths)) %>% pull(), format="f", big.mark=",", digits=0)` total deaths  


Globally, as of the latest WHO data release on `r format(max(global$Date), '%B %d, %Y')`, there are: 

* `r formatC(global %>% group_by(Date_reported) %>% summarize(Total = sum(Cumulative_cases)) %>% ungroup() %>% filter(Date_reported == max(Date_reported)) %>% pull(), format="f", big.mark=",", digits=0)` total cases
  + `r formatC(global %>% group_by(Date_reported) %>% summarize(Total = sum(New_cases)) %>% ungroup() %>% filter(Date_reported == max(Date_reported)) %>% pull(), format="f", big.mark=",", digits=0)` today
* `r formatC(global %>% group_by(Date_reported) %>% summarize(Total = sum(Cumulative_deaths)) %>% ungroup() %>% filter(Date_reported == max(Date_reported)) %>% pull(), format="f", big.mark=",", digits=0)` total deaths
  + `r formatC(global %>% group_by(Date_reported) %>% summarize(Total = sum(New_deaths)) %>% ungroup() %>% filter(Date_reported == max(Date_reported)) %>% pull(), format="f", big.mark=",", digits=0)` today


## Overall Covid Cases In The USA

Regions are defined as follows: Northeast (CT, ME, MA, NH, RI, VT, NJ, NY, PA), Midwest (IN, IL, MI, OH, WI, IA, KS, MN, MO, NE, ND, SD), South (AL, AR, DC, DE, FL, GA, KY, LA, MD, MS, NC, OK, SC, TN, TX, VA, WV) and West (AZ, CO, ID,  NM, MT, UT, NV, WY, AK, CA, HI, OR, WA).  The red line denotes the date that Health and Human Services (HHS) took over the production of data from the Center for Disease Control (CDC) on July 20, 2020.^[See e.g., https://www.cnbc.com/2020/07/20/hhs-unveils-new-coronavirus-hospitalization-database-says-its-more-complete-than-cdcs.html]


```{r chow test, echo=F, eval = F}
## Chow structural break test 
## https://cran.r-project.org/web/packages/strucchange/vignettes/strucchange-intro.pdf

library(tseries)
library(strucchange)

## load the data from a CSV or Excel file. This example is done with an Excel sheet.
cases_per_day <- CDC_w_trendline %>%
  ungroup() %>% group_by(Date) %>%
  summarize(total_new_cases = sum(new_cases))

break_test <- ts(cases_per_day$total_new_cases, start = CDC_w_trendline$Day, frequency = 365)



```


```{r cars, echo = F, warning = F,cache = F, message = F,fig.width = 11, fig.height = 8.5}

New_Cases <- CDC_w_trendline %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Region) %>%
  summarize(new_cases = sum(new_cases)) %>%
  ggplot(aes(x = Date, y = new_cases, fill = Region)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer(direction=-1) +
    geom_vline(xintercept = as.numeric(as.Date("2020-07-20")),
                linetype=4, colour="red") +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Daily Cases") + 
  ggtitle(c(paste("New Cases: CDC Data As Of", as.Date(max(CDC_w_trendline$Date))))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

New_Deaths <- CDC_w_trendline %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Region) %>%
  summarize(new_deaths = sum(new_deaths)) %>%
  ggplot(aes(x = Date, y = new_deaths, fill = Region)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer(direction=-1) +
      geom_vline(xintercept = as.numeric(as.Date("2020-07-20")),
                linetype=4, colour="red") +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Daily Deaths") + 
  ggtitle(c(paste("Deaths Per Day"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

CDC_cases <- CDC_w_trendline %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Region) %>%
  summarise(`Cumulative Cases` = sum(total_cases_by_day),
            total_deaths_by_day = sum(total_deaths_by_day)) %>%
  ggplot(aes(x = Date, y = `Cumulative Cases`, fill = Region)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer(direction=-1) +
      geom_vline(xintercept = as.numeric(as.Date("2020-07-20")),
                linetype=4, colour="red") +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Total Cases") + 
  ggtitle(c(paste("Total Cases")))+ #, as.Date(max(CDC_w_trendline$Date))))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

CDC_deaths <- CDC_w_trendline %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Region) %>%
  summarise(total_cases_by_day = sum(total_cases_by_day),
            `Cumulative Deaths` = sum(total_deaths_by_day)) %>%
  ggplot(aes(x = Date, y = `Cumulative Deaths`, fill = Region)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer(direction=-1) +
      geom_vline(xintercept = as.numeric(as.Date("2020-07-20")),
                linetype=4, colour="red") +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Total Deaths") + 
  ggtitle(c(paste("Total Deaths"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )


covid_tracking_data_w_trendline$Region <- as.character(sapply(covid_tracking_data_w_trendline$State, function(x) names(region.list)[grep(x,region.list)]))

tests_per_day <- covid_tracking_data_w_trendline%>%
  ungroup() %>% #filter(state %in% c("MA","NY") & Date > "2020-07-01") %>%
  group_by(Date, Region) %>%
  filter(state %in% state.abb) %>%
  summarise(tests_by_region = sum(totalTestResults)) %>%
  ungroup() %>% group_by(Region) %>%
  mutate(`New Tests` = tests_by_region -  dplyr::lag(tests_by_region, default = 0)) %>%
  ggplot(aes(x = Date, y = `New Tests`, fill = Region)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer(direction=-1) +
      geom_vline(xintercept = as.numeric(as.Date("2020-07-20")),
                linetype=4, colour="red") +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("New Tests") + 
  ggtitle(c(paste("New Tests: Covidtracker Data as of",max(covid_tracking_data$Date)))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

total_tests <- covid_tracking_data_w_trendline %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Region) %>%
  filter(state %in% state.abb) %>%
  summarise(`Cumulative Tests` = sum(totalTestResults)) %>%
  ggplot(aes(x = Date, y = `Cumulative Tests`, fill = Region)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer(direction=-1) +
      geom_vline(xintercept = as.numeric(as.Date("2020-07-20")),
                linetype=4, colour="red") +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Total Tests") + 
  ggtitle(c(paste("Total Tests"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold")
  )

#CDC
grid.arrange(New_Cases, CDC_cases, New_Deaths, CDC_deaths, tests_per_day, total_tests, nrow = 3, ncol = 2)

# a <- list(
#   text = c(paste("Total Cases To", as.Date(max(CDC_w_trendline$Date)))),
#   xref = "paper",
#   yref = "paper",
#   yanchor = "bottom",
#   xanchor = "center",
#   align = "center",
#   x = 0.5,
#   y = 1,
#   showarrow = FALSE
# )
# 
# b <- list(
#   text = "Total Deaths",
#   xref = "paper",
#   yref = "paper",
#   yanchor = "bottom",
#   xanchor = "center",
#   align = "center",
#   x = 0.5,
#   y = 1,
#   showarrow = FALSE
# )
# 
# subplot(style(ggplotly(CDC_cases)  %>% layout(annotations = a), showlegend = F), 
#         style(ggplotly(CDC_deaths)  %>% layout(annotations = b), showlegend = T),nrows = 1)

```

## US Cases Compared to World Cases
```{r US v World, echo = F, warning = F,cache = F, message = F,fig.width = 11, fig.height = 8.5}

# global <- try_url("https://covid19.who.int/WHO-COVID-19-global-data.csv")
# 
# global$Date <- lubridate::ymd(gsub(pattern = "T.{1,}$", replacement = "", x = global$Date_reported))
global %>% 
    mutate(
         Day = lubridate::yday(Date),
         Week = lubridate::week(Date),
         Country = ifelse(Country_code == "US", "USA", "Rest of World"),
         Country = ifelse(WHO_region == "EURO","Europe", Country), #          Country = ifelse(WHO_region == "EURO","Europe", Country),
         Country = ifelse(is.na(Country), "Rest of World", Country)) %>%
  group_by(Country, Date, Day, Week) %>%
   summarise(Cumulative_cases = sum(Cumulative_cases, na.rm = T),
             New_cases = sum(New_cases, na.rm = T)) %>%
    ggplot(aes(x = Date, y = Cumulative_cases, fill = Country)) +
  geom_area() +
  theme_minimal() +
  scale_fill_brewer() +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Cases") + 
  ggtitle(c(paste("US Cases Compared To World Cases"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold")
  )

```


## Covid Cases in Somerville, MA and MA
```{r Somerville, echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
# from https://somerville-dashboardcovid.trial.opendatasoft.com/explore/dataset/daily-c19-case-counts/information/

MA_and_Somerville <- CDC_w_trendline %>% filter(State == "MA" & Day %in% (max(Day)-14):max(Day)) %>%
  ggplot(aes(x = Date, y = new_cases)) +
  geom_point(size = 3, alpha = 1,col = "darkblue") + 
  geom_line(aes(Date, smooth_new_cases), color="darkblue") +
  geom_point(data = Somerville_MA, mapping = aes(x = Date,y = new_cases), col = 'darkred') +
  geom_line(data = Somerville_MA, mapping = aes(Date, smooth_new_cases), color="darkred") +
  scale_y_continuous(label = scales::number_format(accuracy = 1))  +
  theme_minimal() +
  xlab("Date") + ylab("New Cases") + ggtitle(c(paste("Cases in Somerville, MA (Red) Compared and MA (Blue) as of", max(CDC_w_trendline$Date)))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold")
  )

MA_and_Somerville

# home <- local %>% 
#     mutate(new_cases = Positive - dplyr::lag(Positive, default = 0),
#          new_deaths = Fatalities - dplyr::lag(Fatalities, default = 0),
#          Day = lubridate::yday(Date),
#          Week = lubridate::week(Date)) %>%
#     mutate( smooth_new_cases = ksmooth(Day, new_cases, kernel = "normal", bandwidth = 7)$y) %>%
#   filter(Day %in% c(max(Day), max(Day)-14)) %>%
#   ggplot(aes(x = Date, y = new_cases)) +
#   geom_point(size = 3, alpha = 1) + 
#   geom_line(aes(Date, smooth_new_cases), color="darkblue") +
#   scale_y_continuous(label = scales::number_format(accuracy = 1)) +
#   theme_minimal() + ggtitle(c(paste("Cases Up To", max(local$Date)))) +
#   xlab("Date") + ylab("New Cases Per Day") +
#   theme(
#     plot.title = element_text(size=14, face="bold",hjust = 0.5),
#     axis.title.x = element_text(size=14, face="bold"),
#     axis.title.y = element_text(size=14, face="bold"),
#     strip.text.x = element_text(size=10, face="bold"),
#     axis.text = element_text(size=14, face="bold")
#   )
# 
# MA <- CDC_w_trendline %>% filter(State == "MA" & Week %in% c(max(Week), max(Week)-1)) %>%
#   ggplot(aes(x = Date, y = new_cases)) +
#   geom_point(size = 3, alpha = 1) + 
#   geom_line(aes(Date, smooth_new_cases), color="darkblue") +
#   scale_y_continuous(label = scales::number_format(accuracy = 1))  +
#   theme_minimal() +
#   xlab("Date") + ylab("New Cases Per Day") + ggtitle(c(paste("Cases Up To", max(CDC_w_trendline$Date)))) +
#   theme(
#     plot.title = element_text(size=14, face="bold",hjust = 0.5),
#     axis.title.x = element_text(size=14, face="bold"),
#     axis.title.y = element_text(size=14, face="bold"),
#     strip.text.x = element_text(size=10, face="bold"),
#     axis.text = element_text(size=14, face="bold")
#   )
# 
# grid.arrange(home, MA, nrow = 2, ncol = 2)

```

```{r MA Cases,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
## Covid Cases in MA in the Last Two Weeks 
# New cases by state by day, log scale, with week long trend line
# use log scale otherwhise NY and MA overwhelm other states
```

## Latest Cases in Boston Area (Today, One Week Ago, Two Weeks Ago)

```{r MA Local,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}

Latest_Near_Me <- 
CDC_total %>% filter(countyFIPS %in% c(25017,25021, 25025) & 
                       Week %in% c((max(Week)-3):max(Week) )) %>% 
  ungroup () %>%
  group_by(countyFIPS) %>%
  arrange(Date) %>%
  mutate(`New Cases` = Cases - dplyr::lag(Cases, default = 0),
         `New Deaths` = Deaths - dplyr::lag(Deaths, default = 0)) %>% 
  filter(Date %in% c((max(Date)-14),(max(Date)-7), max(Date) )) %>%
  arrange(`County Name`, desc(Date)) %>% 
  mutate(Date = format(Date, '%B %d, %Y')) %>%
  ungroup() 

knitr::kable(Latest_Near_Me %>% select(Date,`County Name`, Cases, Deaths, `New Cases`, `New Deaths`)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```

## New Covid Cases By State

The plot below shows the number of new COVID cases by state.  Note that each figure has its own y axis; this is done so that trends can be seen within each state.  Log scales can make the difference between 100 and 1000 cases hard to discern and a common y axis causes states like NY and CA to swamp states like NC.  Negative new cases can occur when States revise their total number of cases downwards, as Hawaii did between June 30, 2020 and July 1, 2020.

```{r,echo = F, warning = F,cache = F, message = F, fig.width = 11, fig.height = 17}
# New cases by state by day, log scale, with week long trend line
# use log scale otherwhise NY and MA overwhelm other states
CDC_w_trendline %>% #filter(State == "OK") %>%
  ggplot(aes(x = Date, y = new_cases)) +
  geom_point(size = 1, alpha = .5) + 
  geom_line(aes(Date, smooth_new_cases), color="darkblue") +
  scale_y_continuous(label = scales::number_format(accuracy = 1)) + #,trans='log10') +
  facet_wrap(~ State, nrow = 20 , scales = "free") + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("New Cases Per Day") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10)

  )
```

## COVID Cases per 1,000 Residents by State

This figure shows the number of COVID cases per 1,000.

```{r,echo = F, warning = F,cache = F, message = F, fig.width = 11, fig.height = 17}
# New cases by state by day, log scale, with week long trend line
# use log scale otherwhise NY and MA overwhelm other states
CDC_w_trendline %>% #filter(State == "OK") %>%
  full_join(population_data_by_state, by = "State") %>%
  mutate(cases_per_capita = total_cases_by_day/population * 1000) %>%
  ggplot(aes(x = Date, y = cases_per_capita)) +
  geom_point(size = 1, alpha = .5) + 
  scale_y_continuous(label = scales::number_format(accuracy = 1)) +
  facet_wrap(~ State, nrow = 10) + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("Cases per 1,000") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10)

  )
```


## Latest County Case Map
```{r,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
# Overall level on the most recent day
usmap::plot_usmap("counties", labels = F, include = state.abb,
                  data = CDC_total %>% 
                    filter(countyFIPS > 10 & # & Date == max(Date),
                             State %in% state.abb &
                           Date == max(Date)) %>%
                    dplyr::rename(fips = countyFIPS, 
                                  abbr = State) %>%
                    select(fips, Date, Cases), 
                  values = "Cases") + 
  scale_fill_continuous(low = "blue", high = "red",name = "Covid Cases", label = scales::comma) + #trans = "log10",
  theme(legend.position = "right") + 
  facet_wrap(~ Date)  + 
  theme_void() + xlab("") + ylab("") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=14, face="bold"),
  )

```

## Latest County Death Map
```{r deaths_by_county,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
# Overall level on the most recent day
# Deaths by county
usmap::plot_usmap("counties", labels = F, include = state.abb,
                  data = CDC_total %>% 
                    filter(countyFIPS > 10 & 
                             State %in% state.abb &
                             Date == max(Date)) %>%
                    dplyr::rename(fips = countyFIPS, 
                                  abbr = State) %>%
                    select(fips, Date, Deaths), 
                  values = "Deaths") + 
  scale_fill_continuous(low = "blue", high = "red",name = "COVID Deaths", label = scales::comma) + #trans = "log10",
  theme(legend.position = "right") + 
  theme_void() + xlab("") + ylab("") 
```


## Counties with the Most Cases Over Time
```{r top counties,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}

# Top 5 counties by first of the month
Top5Counties <- CDC_total %>% 
  filter(countyFIPS > 10 & 
           State %in% state.abb &
           (grepl("-01$", Date))|Date == max(Date)) %>%
  dplyr::rename(fips = countyFIPS, 
                abbr = State) %>%
  group_by(fips, Date) %>%
  group_by(Date) %>%
  arrange( -Cases) %>%
  slice(1:5) %>% 
  ungroup() %>% 
  filter(Month %in% c((max(Month)-2):max(Month))) %>%
  arrange(desc(Date))

# See top counties.  Mapping them doesn't make sense

knitr::kable(Top5Counties %>% select(Date, abbr,`County Name`, Cases, Deaths) %>% dplyr::rename(State = abbr)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```


## Hospitalizations By State

```{r hospitalizations,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
covid_tracking_data %>% 
  filter(State %in% state.abb) %>%
  mutate( Day = lubridate::yday(Date),
          Week =  lubridate::week(Date)) %>%
  #select(Date, Day, Week,hospitalizedCurrently) %>% 
  arrange(Date) %>%
  mutate(smooth_hospital = ksmooth(Day, hospitalizedCurrently, kernel = "normal", bandwidth = 7)$y) %>%
  ggplot(aes(x = Date, y = hospitalizedCurrently)) +
  geom_point(size = 1, alpha = .5) + 
  geom_line(aes(Date, smooth_hospital), color="darkblue") +
  scale_y_continuous(label = scales::number_format(accuracy = 1)) +
  facet_wrap(~ State, scales = "free") + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("Currently Hospitalized Patients") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10, face="bold")

  )

```

## Patients Currently in the ICU

```{r icu, echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 10}
covid_tracking_data %>% 
  filter(State %in% state.abb) %>%
  mutate( Day = lubridate::yday(Date),
          Week =  lubridate::week(Date)) %>%
  arrange(Date) %>%
  mutate(smooth_ICU = ksmooth(Day, inIcuCurrently, kernel = "normal", bandwidth = 7)$y) %>%
  ggplot(aes(x = Date, y = inIcuCurrently)) +
  geom_point(size = 1, alpha = .5) + 
  geom_line(aes(Date, smooth_ICU), color="darkblue") +
  scale_y_continuous(label = scales::number_format(accuracy = 1)) +
  facet_wrap(~ State, scales = "free") + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("Currently Hospitalized Patients in the ICU") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10, face="bold")
  )
```

## Patients Currently on Ventilators 

```{r vent, echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 10}
covid_tracking_data %>% 
  filter(State %in% state.abb) %>%
  mutate( Day = lubridate::yday(Date),
          Week =  lubridate::week(Date)) %>%
  arrange(Date) %>%
  mutate(smooth_ICU = ksmooth(Day, onVentilatorCurrently, kernel = "normal", bandwidth = 7)$y) %>%
  ggplot(aes(x = Date, y = onVentilatorCurrently)) +
  geom_point(size = 1, alpha = .5) + 
  geom_line(aes(Date, smooth_ICU), color="darkblue") +
  scale_y_continuous(label = scales::number_format(accuracy = 1)) +
  facet_wrap(~ State, scales = "free") + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("Currently on Ventilator") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10, face="bold")
  )

```


## Cumulative Cases by County in the Northeast
```{r New Northeast, echo = F, warning = F,cache = F, message = F, fig.width = 11, fig.height = 8.5}
#Plot New England on the first of each month
usmap::plot_usmap("counties", labels = F, include = c(.northeast_region), color = "black",
                  data = CDC_total %>%
                    filter(Date > "2020-02-01" & 
                    State %in% .northeast_region &
                    (grepl("-01$", Date)) &
                      countyFIPS > 10 ) %>%
  dplyr::rename(fips = countyFIPS, 
                abbr = State), 
                  values = "Cases") + 
  scale_fill_continuous(low = "blue", high = "red",name = "Covid Cases", label = scales::comma) + #trans = "log10",
  theme(legend.position = "right") + 
  facet_wrap(~ Date)  + 
  theme_void() + xlab("") + ylab("") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=14, face="bold")
  )

```


## Cumulative Cases by State Over Time

```{r,, echo = F, warning = F,cache = F, message = F, fig.width = 11, fig.height = 8.5}
# Cases over time
plot_usmap(data = CDC_total %>% 
             dplyr::rename(state = State) %>%
             filter(Date > "2020-02-01" & 
                      grepl("-01$", Date)) %>%
             group_by(state, Date) %>%
             summarise(Cases = sum(Cases, na.rm = T)), values = "Cases", color = "black") + 
  scale_fill_continuous(low = "blue", high = "red",name = "Covid Cases", label = scales::comma) +
  theme(legend.position = "right") + 
  facet_wrap(~ Date) + 
  theme_void() + xlab("") + ylab("") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=14, face="bold")
  )
```


## Cumulative Deaths by State Over Time

```{r deaths by state,  echo = F, warning = F,cache = F, message = F, fig.width = 11, fig.height = 8.5}
plot_usmap(data = CDC_total %>% 
             dplyr::rename(state = State) %>%
             filter(Date > "2020-02-01" & 
                      grepl("-01$", Date)) %>%
             group_by(state, Date) %>%
             summarise(Deaths = sum(Deaths, na.rm = T)), values = "Deaths", color = "black") + 
  scale_fill_continuous(low = "blue", high = "red",name = "Covid Deaths", label = scales::comma) +
  theme(legend.position = "right") + 
  facet_wrap(~ Date) + 
  theme_void() + xlab("") + ylab("") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=14, face="bold")
    )


```

## Daily Percent Positive Cases 

According to the WHO the positive rate should be below 5%.^[https://coronavirus.jhu.edu/testing/tracker/overview]  Days above the threshold are denoted in red, days below are in blue.  Note that the variables currently used for these plots --positiveIncrease and negativeIncrease-- are being phased out and will be replaced by positiveCasesViral and negativeCasesViral.  Each dot represents the rolling average of tests within the previous week (this approach smoothes out trends such as fewer cases being reported on weekends).

```{r ratio, echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 10}
covid_tracking_data %>% 
  filter(State %in% state.abb) %>%
  mutate( Day = lubridate::yday(Date),
          Week =  lubridate::week(Date),
          Pos_roll_avg = zoo::rollmean(positiveIncrease, k = 7, fill = NA),
          Neg_roll_avg = zoo::rollmean(negativeIncrease, k = 7, fill = NA),
          ratio = Pos_roll_avg/(Pos_roll_avg + Neg_roll_avg)) %>%
  arrange(Date) %>%
  filter(Week %in% c((max(Week)-3):max(Week))) %>%
  ggplot(aes(x = Date, y = ratio)) +
  geom_point(aes(colour = cut(ratio, c(-Inf, 0.05, Inf))),
             size = 1) +
    scale_color_manual(name = "ratio",
                     values = c("(-Inf,0.05]" = "darkblue",
                                "(0.05, Inf]" = "red"),
                     labels = c("Falling","Rising")) +
  #geom_line(aes(Date, smooth_ratio), color="darkblue") +
  scale_y_continuous(label = scales::percent_format(accuracy = 0.1)) +
  facet_wrap(~ State) + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("Ratio of Positive to Negative Tests") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10, face="bold"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    legend.position = "none"
  )

```

## Partisanship and COVID Cases

```{r partisan, echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 10}
load("/Users/stevenrashin/Documents/GitHub/Covid-Tracker/MIT_elections_presidential_returns.rdata")

election_2016 <- x %>% filter(year == 2016 & candidate %in% c("Hillary Clinton","Donald Trump")) %>%
  rename(countyFIPS = FIPS)

county_election_2016 <- election_2016 %>% as.tibble() %>%
  group_by(countyFIPS) %>%
  summarise(DemVotes = ifelse(party == "democrat",candidatevotes,0),
            RepVotes = ifelse(party == "republican",candidatevotes,0)) %>%
  summarise(DemVotes = sum(DemVotes),
            RepVotes = sum(RepVotes)) %>% group_by(countyFIPS) %>%
  summarise(Winner = ifelse(DemVotes > RepVotes,"Dem","Rep"))

state_election_2016 <- election_2016 %>% group_by(state_po, candidate) %>% 
  summarize(votes = sum(candidatevotes,na.rm = T)) %>%
  spread(candidate, votes, fill = NA, convert = FALSE) %>%
  mutate(Winner = ifelse(`Hillary Clinton`>`Donald Trump`,"Dem","Rep")) %>% 
  rename(State = state_po) %>% select(State, Winner) %>%
  ungroup()
  
state_level_potus <- CDC_w_trendline %>% 
  full_join(state_election_2016, by = "State") %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Winner) %>%
  summarize(new_cases = sum(new_cases)) %>%
  ggplot(aes(x = Date, y = new_cases, fill = Winner)) +
  geom_area() +
  theme_minimal() +
  scale_fill_manual(values = alpha(c("blue", "red"), .7)) +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Daily Cases") + 
  ggtitle(c(paste("Daily Cases by 2016 State Level Presidential Election Result"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

county_level_potus <- CDC_total %>% filter(countyFIPS > 10) %>%
  full_join(county_election_2016, by = "countyFIPS") %>% 
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(countyFIPS) %>%
  mutate(new_cases = Cases - dplyr::lag(Cases, default = 0)) %>%
  group_by(Date, Winner) %>%
  summarise(new_cases = sum(new_cases)) %>%
  ggplot(aes(x = Date, y = new_cases, fill = Winner)) +
  geom_area() +
  theme_minimal() +
  scale_fill_manual(values = alpha(c("blue", "red"), .7)) +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Daily Cases") + 
  ggtitle(c(paste("By County"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

State_Parties <- read.csv("/Users/stevenrashin/Documents/GitHub/Covid-Tracker/state_parties.csv", skip = 3, stringsAsFactors = F)
State_Parties$State <- state.abb[match(State_Parties$Location,state.name)]
State_Parties <- State_Parties %>% filter(State %in% state.abb)

state_level_gov <- CDC_w_trendline %>% 
  full_join(State_Parties, by = "State") %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date, Governor.Political.Affiliation) %>%
  summarize(new_cases = sum(new_cases)) %>%
  ggplot(aes(x = Date, y = new_cases, fill = Governor.Political.Affiliation)) +
  geom_area() +
  theme_minimal() +
  scale_fill_manual(values = alpha(c("blue", "red"), .7)) +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Daily Cases") + 
  ggtitle(c(paste("Daily Cases by Partisanship of Current Governor"))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold"),
    legend.position = "none"
  )

#### Show Data is Different ####
grid.arrange(state_level_potus, county_level_potus, state_level_gov, nrow = 1)




```

# Variants of Select Figures

## Week to Week Changes in COVID Cases

```{r week-to-week,echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
# Week-to-week percentage change in cases: (new week - prior week) / prior week
good_or_bad_week <- CDC_w_trendline  %>% 
  group_by(State, Week) %>%
  dplyr::summarise(new_cases_per_week = sum(new_cases, na.rm = T)) %>%
  mutate(
    last_week = lag(new_cases_per_week, 1),
    pct_change = (new_cases_per_week - last_week)/last_week , 
    week_to_date = lubridate::ymd( "2020-01-01" ) + lubridate::weeks(Week - 1 ),
    moving_average = zoo::rollmean(new_cases_per_week,4,align='right',fill=NA),
    pct_change_month = (new_cases_per_week - moving_average)/moving_average) %>%   
  #replace(is.na(.), 0)  %>% 
  group_by(State) %>%
  filter(Week > 10) 

good_or_bad_week %>% mutate(alpha1 = 1/Week) %>%
  ggplot(aes(x = week_to_date, y = pct_change)) + 
  geom_point(aes(colour = cut(pct_change, c(-Inf, 0, Inf))),
             size = 1) +
  scale_color_manual(name = "Cases",
                     values = c("(-Inf,0]" = "darkblue",
                                "(0, Inf]" = "red"),
                     labels = c("Falling","Rising")) +
  scale_y_continuous(label = scales::percent_format(accuracy = 1)) +
  facet_wrap(~ State) + #, scales = "free") + # free scales scales each state individually
  theme_minimal() +  #geom_line(alpha = 0.3) +
  xlab("Date") + ylab("Change in New Cases from Prior Week") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10, face="bold"),
    legend.position = "none"
  ) +
  coord_cartesian(ylim = c(-0.50,0.50)) +
  geom_hline(yintercept=0, color = "black", linetype = "dashed", size=0.5) +
  scale_alpha_continuous(guide=FALSE)

```

## Confirmed Cases as a Percentage of State Population

The two sets of points are confirmed cases (black), and potential cases (grey) which are confirmed cases times ten.  For herd immunity, typically 70-90% of a population needs immunity.^[See <https://www.jhsph.edu/covid-19/articles/achieving-herd-immunity-with-covid19.html>]

```{r, echo = F, warning = F,cache = F, message = F, fig.width = 13, fig.height = 8.5}
# Total cases by state, as a percentage of state population
CDC_w_trendline %>% #filter(State == "AK") %>%
  left_join(population_data_by_state, by = "State") %>%
  mutate(cases_as_pct_of_pop = total_cases_by_day/population) %>%
  ggplot(aes(x = Date, y = cases_as_pct_of_pop)) +
  geom_point(size = 1, alpha = .5) + #color = "grey"
  geom_point(mapping = aes(x = Date, y = cases_as_pct_of_pop * 10), size = 1, alpha = .3, color = "grey") + #color = "grey"
  #geom_line(aes(Date, smooth_total), color="darkblue") +
  scale_y_continuous(label = scales::percent_format(accuracy = 0.1)) +
  facet_wrap(~ State) + # free scales scales each state individually
  theme_minimal() +
  xlab("Date") + ylab("Cases as % of State Population") +
  theme(
    plot.title = element_text(size=14, face="bold"),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=10, face="bold")

  )

```


## Data Sources

Data for this report comes from:

* The CDC Homepage <https://www.cdc.gov/covid-data-tracker/index.html#county-map> and <https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/>, specifically:
    + Confirmed COVID Cases <https://usafactsstatic.blob.core.windows.net/public/data/covid-19/covid_confirmed_usafacts.csv>
    + Covid Deaths <https://usafactsstatic.blob.core.windows.net/public/data/covid-19/covid_deaths_usafacts.csv>
    + County Population <https://usafactsstatic.blob.core.windows.net/public/data/covid-19/covid_county_population_usafacts.csv>
* Covidtracker.com data (used by Johns Hopkins <https://coronavirus.jhu.edu/us-map>) <https://covidtracking.com/api/v1/states/daily.csv>
* World Health Organization <https://covid19.who.int/>
* Somerville City data <https://somerville-dashboardcovid.trial.opendatasoft.com/pages/citysdashboard/>


## Different Counts from Different Data

See the difference between <https://covidtracking.com/api/v1/states/daily.csv> and <https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/>

```{r differences, echo = F, warning = F,cache = F, message = F, fig.width = 11, fig.height = 8.5}
covid_tracker <- covid_tracking_data_w_trendline  %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date) %>%
  replace(is.na(.), 0) %>%
  summarise(total_cases_by_day = sum(positive, na.rm = T),
            total_deaths_by_day = sum(death, na.rm = T)) %>%
  ggplot(aes(x = Date, y = total_cases_by_day)) +
  geom_point(size = 1, alpha = .5) + 
  geom_point(mapping = aes(x = Date, y = total_deaths_by_day),
             size = 1, alpha = .5, shape = 2, color = "darkred") +
  theme_minimal() +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Total Cases/Deaths") + 
  ggtitle(c(paste("Covidtracking.com Through",max(covid_tracking_data_w_trendline$Date)))) + 
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold")

  )

### Overall Cases in USA ###
CDC <- CDC_w_trendline %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date) %>%
  summarise(total_cases_by_day = sum(total_cases_by_day),
            total_deaths_by_day = sum(total_deaths_by_day)) %>%
  ggplot(aes(x = Date, y = total_cases_by_day)) +
  geom_point(size = 1, alpha = .5) +
  geom_point(mapping = aes(x = Date, y = total_deaths_by_day),
             size = 1, alpha = .5, shape = 2, color = "darkred") +
  theme_minimal() +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Total Cases/Deaths") +
  ggtitle(c(paste("CDC Through", as.Date(max(CDC_w_trendline$Date))))) +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold")
  )


merge_covid_tracker <- covid_tracking_data_w_trendline  %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date) %>%
  replace(is.na(.), 0) %>%
  summarise(CT_total_cases_by_day = sum(positive, na.rm = T),
            CT_total_deaths_by_day = sum(death, na.rm = T))

merge_CDC <- CDC_w_trendline  %>%
  ungroup() %>% #filter(State %in% c("MA","NY") & Week > 20) %>%
  group_by(Date) %>%
  replace(is.na(.), 0) %>%
  summarise(CDC_total_cases_by_day = sum(total_cases_by_day, na.rm = T),
            CDC_total_deaths_by_day = sum(total_deaths_by_day, na.rm = T))

difference_in_counts <- merge(merge_CDC, merge_covid_tracker, by = c("Date"))

difference_in_counts <- difference_in_counts %>% 
  select(Date, CDC_total_cases_by_day, CT_total_cases_by_day, 
         CDC_total_deaths_by_day, CT_total_deaths_by_day) %>%
  mutate(case_difference = CDC_total_cases_by_day - CT_total_cases_by_day, 
         death_difference = CDC_total_deaths_by_day - CT_total_deaths_by_day)

difference_plot <- difference_in_counts %>%
  ggplot(aes(x = Date, y = case_difference)) +
  geom_point(size = 1, alpha = .5) + 
  geom_point(mapping = aes(x = Date, y = death_difference),
             size = 1, alpha = .5, shape = 2, color = "darkred") +
  theme_minimal() +
  scale_y_continuous(label = scales::comma) +
  xlab("Date") + ylab("Difference in US Cases/Deaths") + 
  ggtitle("CDC Minus Covidtracking.com") +
  theme(
    plot.title = element_text(size=14, face="bold",hjust = 0.5),
    axis.title.x = element_text(size=14, face="bold"),
    axis.title.y = element_text(size=14, face="bold"),
    strip.text.x = element_text(size=10, face="bold"),
    axis.text = element_text(size=14, face="bold")
  )

#### Show Data is Different ####
grid.arrange(CDC, covid_tracker, difference_plot, nrow = 1)

```



